---
title: "STAT 508 Final Project EDA"
author: "Callum Arnold"
output:
    html_notebook:
        code_folding: hide
        toc: yes
        toc_float: yes
---

# Set Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 14)
```

```{r}
library(tidyverse)
library(here)
library(hrbrthemes)
library(janitor)
library(corrplot)
library(skimr)

RNGkind(sample.kind = "Rounding")
set.seed(1)

theme_set(theme_ipsum())
```

```{r}
credit <- as_tibble(read_csv(here("data", "creditcard.csv")))
```

# EDA

```{r}
head(credit)
```

## Variable Summary

```{r}
anyNA(credit)
```

```{r}
skim(credit)
```


```{r}
for (i in 1:ncol(credit)){
  print(summary(credit[, i]))
}
```

### Distributions

```{r}
for (i in names(credit[, -31])) {
  p <- ggplot(credit, aes_string(x = i)) +
    geom_density(fill = "cornsilk")

  print(p)
}
```

## Relationship with Response

```{r}
ggplot(data = credit, aes(x = Time, fill = factor(Class))) +
  geom_histogram() +
  facet_wrap(~Class, scales = "free")
```

```{r}
ggplot(data = credit, aes(x = log(Amount), fill = factor(Class))) +
  geom_histogram() +
  facet_wrap(~Class, scales = "free")
```

```{r}
ggplot(data = credit, aes(x = Time, y = log(Amount), alpha = 0.2, color = factor(Class))) +
  geom_point() +
  facet_wrap(~Class, scales = "free")
```
There is a clear skew to the `Amount` variable, so it is worth applying a transformation to the data. 
As there are values of `0`, we need to add `1` to ensure that we don't get `Inf` values produced after log transformation.

```{r}
credit <- credit %>%
  mutate(log_amount = log(Amount + 1))

summary(credit$log_amount)
```

```{r}
ggplot(credit, aes(x = log_amount)) +
    geom_density(fill = "cornsilk")
```

## Predictor Correlations

```{r}
corrplot(cor(credit[, -31]), method = "square", type = "upper")
```

Transformation of `Amount` has helped to reduce collinearities of the predictors.
The only correlations are between `V3` and `Time`, and `V2` and `log_amount`.
This is as expected as PCA produces orthogonal linear combinations,
therefore there shouldn't be much correlation between them.

```{r}
credit %>%
  mutate(
    trans_time = scale(Time, center = TRUE),
    trans_amount = scale(log_amount, center = TRUE)
    ) %>%
  select(trans_time, everything()) %>%
  select(-c(Amount, Time, log_amount)) %>%
  cor() %>%
  corrplot(method = "square", type = "upper")
```

# PCA Importance

```{r}
vars <- credit %>%
  select(-c(Time, Amount, log_amount, Class)) %>%
  sapply(var) %>%
  sort(decreasing = TRUE) %>%
  as_tibble()

vars
```

```{r}
prop_vars <- vars %>%
  transmute(variance = value) %>%
  mutate(prop_var = variance / sum(variance),
         cum_var = cumsum(prop_var),
         PC = factor(paste0("V", row_number()), levels = paste0("V", row_number())))

levels(prop_vars$PC)

prop_vars %>%
  ggplot(aes(x = PC, y = prop_var)) +
  geom_col()
```

```{r}
prop_vars %>%
  ggplot(aes(x = PC, y = cum_var)) +
  geom_col()
```



